Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-10 shanghai.tex
I modified and then promoted your remark (at the end of the section introducing CS(G)) to the textual mainstream because it will used in the section on base change over finite fields.

Character sheaves on $G$ form a category, denoted by $\CS(G)$. In
fact, $\CS(G)$ is a rigid symmetric monoidal category, so isomorphism
classes of character sheaves on $G$ form a group, denoted by
$\CSiso{G}$; this is the Picard group of the category of
character sheaves on $G$. Our application of character sheaves to
characters of algebraic tori over non-Archimedean local fields rests
on the Picard group of character sheaves on $G$, for all smooth
commutative group schemes $G$, locally of finite type over $\Fq$, 
for which the geometric component group is finitely generated.

@@

We fix an algebraic closure $\bFq$ of $\Fq$ and write $\bG$ for the
commutative group scheme $G \times_{\Spec{\Fq}} \Spec{\bFq}$, over $\bFq$ obtained by base change from $k$. The multiplication morphism for $\bG$ will be denoted by $\bm$. 
%
Let $\Frob{G}$ be the Frobenius automorphism of $\bG$.
\Clifton{I was a bit glib about Frobenius, previously.
Now I've made things more precise here; I'm in the process of tracking this improvement through the paper.}
Recall that $\Frob{G} = \id_{G}\times \Frob{\Spec{\bFq}}$ where
$\Frob{\Spec{\bFq}} = \Spec\Frob{\Fq}$ and $\Frob{\Fq}$  is the geometric Frobenius element in $\Weil{\Fq} \subset \Gal(\bFq/\Fq)$.

The category of character sheaves on $G$, denoted by \cdef{$\CS(G)$},
is defined as follows.  A \cdef{character sheaf} on $G$ is a triple
$\cs{L}\ceq (\gcs{L},\mu,\phi)$ where:
\begin{enumerate}
\labitem{(CS.0)}{CS.0} $\gcs{L}$ is a rank-$1$ $\ell$-adic
  local system on (the \'etale site of) $\bG$;
\labitem{(CS.1)}{CS.1} $\mu: \bm^* \gcs{L} \to \LxL$
  is an isomorphism of sheaves on $\bGxG$ such that the following diagram commutes, 
  where $m_3 \ceq m\circ (m\tight{1}{\times}{2}\id) = m\circ (\id\tight{2}{\times}{1} m)$;
%
  \[
  \begin{tikzcd}[row sep=30]
  \bm_3^*\gcs{L} \arrow{rr}{(\bm \tight{1}{\times}{2} \id)^*\mu} \arrow[swap]{d}{(\id \tight{2}{\times}{1} \bm)^*\mu}
    &&  \bm^*\gcs{L} \boxtimes \gcs{L} \dar{\mu \tight{0}{\boxtimes}{1} \id} \\
    \gcs{L} \boxtimes \bm^* \gcs{L} \arrow{rr}{\id \boxtimes \mu}
    &&  \gcs{L} \tight{0}{\boxtimes}{0} \LxL\\
  \end{tikzcd}
  \]

%
\labitem{(CS.2)}{CS.2} $\phi : \Frob{G}^* \gcs{L} \to \gcs{L}$ is an
  isomorphism of sheaves on $\bG$ compatible with $\mu$ in the sense that the following diagram commutes.
%
  \[
  \begin{tikzcd}[row sep=20]
  \Frob{\GxxG}^* \bm^* \gcs{L} \arrow{rr}{\Frob{\GxxG}^*\mu}
    && \Frob{\GxxG}^*(\LxL)\\
    \arrow[equal]{u} \bm^*  \Frob{G}^* \gcs{L} \arrow[swap]{d}{\bm^* \phi}
    && \Frob{G}^*\gcs{L}\boxtimes \Frob{G}^*\gcs{L} \dar{F\boxtimes \phi} \arrow[equal]{u} \\
    \bm^*\gcs{L} \arrow{rr}{\mu}
    && \LxL\\
  \end{tikzcd}
  \]
\end{enumerate}
Morphisms in $\CS(G)$ are defined in the natural way:
\begin{enumerate}
\labitem{(CS.3)}{CS.3} if $\cs{L} = (\gcs{L},\mu,\phi)$ and
  $\cs{L'} = (\gcs{L'},\mu',\phi')$ are character sheaves on $G$ then
  $\Hom_{\CS(G)}(\cs{L},\cs{L'})$ is the set of morphisms $\varphi : \gcs{L} \to \gcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that the following diagrams both commute.
  \[
  \begin{tikzcd}[column sep=40]
  \Frob{G}^* \gcs{L} \rar{\Frob{G}^* \varphi} \arrow[swap]{d}{F} & \Frob{G}^* \gcs{L'} \dar{F'}
  & \arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}} & \arrow[swap]{d}{\mu} m^* \gcs{L} \rar{m^* \varphi} & m^* \gcs{L'} \dar{\mu'} \\
  \gcs{L} \rar{\varphi} & \gcs{L'}
  & {} & \LxL \rar{\tight{1}{\varphi\boxtimes \varphi}{1}} & \gcs{L'} \tight{0}{\boxtimes}{0} \gcs{L'}
  \end{tikzcd}
  \]
\end{enumerate}

@@

%\begin{proposition}\label{tensor}
The category $\CS(G)$ of character sheaves on $G$ is a rigid monoidal
category [insert reference] under
$\cs{L} \otimes \cs{L'} \ceq (\gcs{L}\otimes\gcs{L'}, \mu\otimes\mu', \phi\otimes \phi')$
with duals given by applying the functor 
\Clifton{Should this be $\sheafHom(\ - \ ,\EE)$?} $\sheafHom(\ - \ ,(\EE)_{\bG})$ (sheaf hom) to
conditions \ref{CS.0}, \ref{CS.1} and \ref{CS.2}.
%\end{proposition}
(The category of characters sheaves on $G$ is not abelian, so $\CS(G)$
is not a tensor category in the sense of \cite{Deligne:tensorielles}.)
%\begin{proof}\end{proof}
%
This rigid monoidal category structure for $\CS(G)$ puts a group
structure on the set $\CSiso{G}$ of isomorphism classes in
$\CS(G)$; we refer to this as the \cdef{Picard group of character
sheaves} on $G$.\footnote{See {\tt http://ncatlab.org/nlab/show/Picard+group}.}

\begin{remark} \label{rem:CS-action}
If $W$ is a group equipped with a homomorphism $\alpha : W \to \Gal(\bFq/\Fq)$
then an \emph{action} of $W$ on $\gcs{L}$ is a collection of isomorphisms
\[
\gamma_w : \alpha(w)^* \gcs{L} \to \gcs{L}
\]
so that $\gamma_{uw} = \gamma_w \circ \alpha(w)^*\gamma_u$. \David{make sure I have the order right and add a reference}
Specifying an $F$ as in \ref{CS.2} is the same as giving an action of the Weil group $\Weil{k}$
on $\gcs{L}$ compatible with $\mu$.
\end{remark}

@@

To define the category of finite character sheaves we must work a bit harder.
Consider the following category $C(G)$:
an object in $C(G)$ is a pair $(f,\psi)$ where: 
$f : H\to G$ is an isogeny (a finite, surjective, etale, morphism of commutative group $\Fq$-schemes) 
for which the action of $\Gal(\bFq/\Fq)$ on the etale group scheme $\ker f$ is trivial, 
and $\psi : \ker f\to \Aut(V)$ is a representation on a finite-dimensional $\EE$-vector space $V$;
a map $(f,\psi) \to (f',\psi')$ in $C(G)$ is a pair $(\varphi,T)$ 
where $\varphi$ is a morphism of schemes such that $f' = f\circ \varphi$
and $T : V\to V'$ is a linear transformation equivariant 
for the action of $\ker f'$ on $V'$ by $\psi'$ 
and the action of $\ker f'$ on $V$ by $\psi \circ \varphi$.

@@

We say that $(\varphi,T)$ in $C(G)$ is a weak isomorphism if $T$ is an isomorphism of vector spaces. 
Let $C(G)[W^{-1}]$ be the localization of $C(G)$ at weak isomorphisms.
Now, let $C_1(G)$ be the subcategory of $C(G)$ such that $\psi$ is a one-dimensional representation
and let $\CSf(G)$ be the localisation of $C_1(G)$ at weak isomorphisms. 
This is the rigid symmetric monoidal category of \cdef{finite character sheaves} on $G$.

@@

\begin{proof}
 To prove the proposition we exhibit a full and faithful monoidal functor $L_G : \CSf(G) \to \CSb(G)$.
 %
 We will do this by constructing a faithful monoidal functor $L : C_1(G) \to \CSb(G)$ 
 such that $L(\varphi,T)$ is an isomorphism precisely when $(\varphi,T)$ is a weak isomorphism. 
 Factoring along $C_1(G) \to \CSb(G)$ will produce the full and faithful monoidal functor we seek.

 Now we introduce a faithful functor $L : C(G) \to \Loc(G)$ to the category of local systems on $G$.
 For any object $(f,\psi)$ in $C(G)$, let $L(f,\psi)$ be the $\psi$-isotypic component of $f_* V_H$, 
 where $V_H$ refers to $V$ as a constant sheaf on $H$. 
 We may identify the stalk $(f_* V_H)_{\bar g}$ of $f_* V_H$ at a geometric point ${\bar g}$ on $G$ 
 with the $\EE$-vector space $\Hom_\text{set}(\abs{f^{-1}({\bar g})},V)$ 
 and the stalk $L(f,\psi)_{\bar g}$ of $L(f,\psi)$ at a geometric point ${\bar g}$ on $G$ 
 with the $\EE$-vector space consisting of those functions $s: \abs{f^{-1}({\bar g})} \to V$ such that 
 $s(a\cdot {\bar h}) = \psi(a)(s({\bar h}))$ for all $a\in \ker f$ and for all ${\bar h} \in f^{-1}({\bar g})$.
 Since $L(f,\psi)$ is trivialized by $f$, it is a local system on $G$.
 For any map $(\varphi,T)$ in $C(G)$, let $L(\varphi,T) : L(f,\psi) \to L(f',\psi')$ be the homomorphism of local systems defined 
 %at geometric points $L(\varphi,T)_{\bar g} : L(f,\psi)_{\bar g} \to L(f',\psi')_{\bar g}$ 
 by $s \mapsto s' \ceq T\circ s\circ \varphi$. 
 This defines a monoidal functor $L$ from $C(G)$ to $\Loc(G)$. 
 This functor commutes with pull-back along any morphism of group schemes.

 Now consider the localization $C(G)[W^{-1}]$ of $C(G)$ at weak isomorphisms.
 Morphisms in $C(G)[W^{-1}]$ are simple to describe because category $C(G)$ 
 admits a calculus of left fractions\footnote{See {\tt http://ncatlab.org/nlab/show/calculus+of+fractions}.} for weak isomorphisms: 
 every isomorphism in $C(G)$ is a weak isomorphism; 
 push-outs exist in $C(G)$ and the push-out of a weak isomorphism is a weak isomorphism; 
 and weak isomorphisms enjoy the left-cancellation property. 
 It follows that a morphism in $C(G)[W^{-1}]$ is the equivalence class of 
  \[
   \begin{tikzcd}
   (f,\psi) \arrow{r}{(\varphi_1,T_1)} & (f_1,\psi_1) & \arrow[swap]{l}{(\varphi'_1,T'_1)} (f',\psi')
   \end{tikzcd}
  \]
 where $(\varphi',T') : (f,\psi) \to (f',T')$ is any map in $C(G)$ 
 and $(\varphi'_1,T'_1) : (f_1,\psi_1) \to (f',\psi')$ is a weak isomorphism; 
 this morphism is commonly denoted by $(\varphi_1,T_1)^{-1}\circ (\varphi'_1,T'_1)$. 
 %The equivalence relation identifies $(\varphi_1,T_1)^{-1}\circ (\varphi'_1,T'_1)$ 
 % with $(\varphi_2,T_2)^{-1}\circ (\varphi'_2,T'_2)$ if there are maps 
 %$(f_1,\psi_1)\to (f_0,\psi_0)$ and $(f_2,\psi_2)\to (f_0,\psi_0)$ making the following diagram commute
 %\[
 %  \begin{tikzcd}
 %  & (f_1,\psi_1) \arrow{d} & \\
 %  (f,\psi) \arrow{ur}{(\varphi_1,T_1)} \arrow{dr}{(\varphi_2,T_2)} & (f_0,\psi_0)  & \arrow[swap]{ld}{(\varphi'_2,T'_2)} \arrow[swap]{lu}{(\varphi'_1,T'_1)} (f',\psi')\\
 %  & (f_2,\psi_2) \arrow{u} & \\
 %  \end{tikzcd}
 %\]
 %where $(f',\psi') \to (f_0,\psi_0)$ is a weak isomorphism.

 Observe that $L$ takes weak isomorphisms in $C(G)$ to isomorphisms in $\Loc(G)$. 
 To see this, let $(\varphi,T) : (f,\psi)\to (f',\psi')$ be a weak isomorphism, 
 where $f:H\to G$ and $f':H'\to G$ as above. 
 Since $T$ is an isomorphism, we may replace $V'$ with $V$ and $T$ with $\id_V$, without loss of generality. 
 Then the equivariance condition on $T$ becomes $\psi' = \psi\circ \varphi$ on the finite etale group scheme $\ker f'$. 
 Since $f' = f \circ \varphi$ and $f$ and $f'$ are etale covers, it follows that $\varphi$ is also an etale cover. 
 It follows that $\varphi : H'\to H$ admits a section $\varphi' : H \to H'$, 
 which provides the inverse to $L(\varphi,T) : L(f,\psi) \to L(f',\psi')$, above. 
 It is also clear that if $L(\varphi,T)$ is an isomorphism of local systems, 
 then $T$ must be an isomorphism of vector spaces, 
 whence $(\varphi,T)$ is a weak isomorphism: $W$ is saturated.

@@

  To produce the isomorphism between $\CSiso{G(\bFq)}$ and
  $\Hh^1(\Weil{\Fq}, G(\bFq)^*)$ we must pick a global section of the line
  bundle $(\gcs{L},\mu)$, for each character sheaf $\cs{L}$ on $G$; 
  such a section exists by Lemma~\ref{lemma:section}.
  We will show that the group isomorphism does not depend on these
  choices.
%
  We use the section $s_\cs{L}$ to define a cocycle $z_\cs{L} \in Z^1(\Weil{\Fq},G(\bFq)^*)$
  as follows. Recall that $\cs{L}$ also determines a family of linear transformations
  $\phi_x : \gcs{L}_{\Frob{\Fq}(x)} \to \gcs{L}_x$ of one-dimensional $\EE$-vector
  spaces, so $\phi_x(s(\Frob{\Fq}(x)))$ is a scalar multiply of $s(x)$, for
  each $x\in G(\bFq)$; let $\tau_\cs{L}(\Frob{\Fq})(x)$ be that scalar, so
  \begin{equation}\label{t}
    \forall x\in G(\bFq),\qquad F_{x}(s(\Frob{\Fq}(x)) = \tau_\cs{L}(\Frob{\Fq})(x) s(x).
  \end{equation}
  Conditions \ref{cs.0} and \ref{cs.2}, together with \eqref{section}, guarantee that
  the scalar $z_\cs{L}(\Frob)(x)$ is non-zero, while condition~\ref{cs.1} forces
  $\tau_\cs{L}(\Frob{\Fq})(x+y) = \tau_\cs{L}(\Frob{\Fq})(x) \ \tau_\cs{L}(\Frob{\Fq})(y)$,
  for all $x,y \in G(\bFq)$.
  % \Clifton{The alternative to this approach is to change the
  %   definition of character sheaves to incorporate the notion of
  %   Galois sheaves. That might not be so bad, but would add another
  %   layer of complication, which we wouldn't use profitably, expect
  %   here.}

@@

\begin{lemma}
 If $\cs{L}$ is a character sheaf then there is a non-trivial section of $\gcs{L}$ 
 compatible with $\mu$ in the sense that 
 \begin{equation}\label{section}
  \forall x,y \in G(\bFq), \qquad \mu_{x,y}(s(x+y)) = s(x)\otimes s(y).
 \end{equation}
\end{lemma}

@@

\bib{SGA4.5}{book}{
   author={Deligne, P.},
   title={Cohomologie \'etale},
   series={Lecture Notes in Mathematics, Vol. 569},
   note={S\'eminaire de G\'eom\'etrie Alg\'ebrique du Bois-Marie 1963-64 SGA
   4$\frac{1}{2}$; 
   Avec la collaboration de J. F. Boutot, A. Grothendieck, L. Illusie et J.
   L. Verdier},
   publisher={Springer-Verlag},
   place={Berlin},
%date={4h},
   date={1977},
   pages={iv+312pp},
%   review={\MR{0463174 (57 \#3132)}},
}

\bib{SGA7.2}{book}{
   title={Groupes de monodromie en g\'eom\'etrie alg\'ebrique. II},
%   language={French},
 author={Deligne, P.},
 author={Katz, N.},
   series={Lecture Notes in Mathematics, Vol. 340},
   note={S\'eminaire de G\'eom\'etrie Alg\'ebrique du Bois-Marie 1967--1969
   (SGA 7 II);
   Dirig\'e par P. Deligne et N. Katz},
   publisher={Springer-Verlag},
   place={Berlin},
   date={1973},
   pages={x+438},
%   review={\MR{0354657 (50 \#7135)}},
}

>>>>

Character sheaves on $G$ form a category, denoted by $\CS(G)$. In
fact, $\CS(G)$ is a rigid symmetric monoidal category, so isomorphism
classes of character sheaves on $G$ form a group, denoted by
$\CSiso{G}$; this is the Picard group of the category of
character sheaves on $G$. Our application of character sheaves to
characters of algebraic tori over non-Archimedean local fields rests
on finding the Picard group of character sheaves on $G$, for all smooth commutative group schemes $G$, locally of finite type over $\Fq$, for which the geometric component group is finitely generated.

@@

We fix an algebraic closure $\bFq$ of $\Fq$ and write $\bG$ for the
commutative group scheme $G \times_{\Spec{\Fq}} \Spec{\bFq}$ over $\bFq$ obtained by base change from $k$. The multiplication morphism for $\bG$ will be denoted by $\bm$.

Let $\Frob{G}$ be the Frobenius automorphism of $\bG$.
Recall that $\Frob{G} = \id_{G}\times \Frob{\Spec{\bFq}}$ where
$\Frob{\Spec{\bFq}} = \Spec{\Frob{\Fq}}$ and $\Frob{\Fq}$  is the geometric Frobenius element in $\Gal(\bFq/\Fq)$.

The category of character sheaves on $G$, denoted by \cdef{$\CS(G)$},
is defined as follows.  A \cdef{character sheaf} on $G$ is a triple
$\cs{L}\ceq (\gcs{L},\mu,\phi)$ where:
\begin{enumerate}
\labitem{(CS.0)}{CS.0} $\gcs{L}$ is a constructible $\ell$-adic sheaf on  (the \'etale site of) $\bG$, locally constant of rank-$1$ on each connected component of $\bG$;
\labitem{(CS.1)}{CS.1} $\mu: \bm^* \gcs{L} \to \LxL$
  is an isomorphism of sheaves on $\bGxG$ such that the following diagram commutes, 
  where $m_3 \ceq m\circ (m\tight{1}{\times}{2}\id) = m\circ (\id\tight{2}{\times}{1} m)$;
%
  \[
  \begin{tikzcd}[row sep=30]
  \bm_3^*\gcs{L} \arrow{rr}{(\bm \tight{1}{\times}{2} \id)^*\mu} \arrow[swap]{d}{(\id \tight{2}{\times}{1} \bm)^*\mu}
    &&  \bm^*\gcs{L} \boxtimes \gcs{L} \dar{\mu \tight{0}{\boxtimes}{1} \id} \\
    \gcs{L} \boxtimes \bm^* \gcs{L} \arrow{rr}{\id \boxtimes \mu}
    &&  \gcs{L} \tight{0}{\boxtimes}{0} \LxL\\
  \end{tikzcd}
  \]

%
\labitem{(CS.2)}{CS.2} $\phi : \Frob{G}^* \gcs{L} \to \gcs{L}$ is an
  isomorphism of constructible $\ell$-adic sheaves on $\bG$ compatible with $\mu$ in the sense that the following diagram commutes.
%
  \[
  \begin{tikzcd}[row sep=20]
  \Frob{\GxxG}^* \bm^* \gcs{L} \arrow{rr}{\Frob{\GxxG}^*\mu}
    && \Frob{\GxxG}^*(\LxL)\\
    \arrow[equal]{u} \bm^*  \Frob{G}^* \gcs{L} \arrow[swap]{d}{\bm^* \phi}
    && \Frob{G}^*\gcs{L}\boxtimes \Frob{G}^*\gcs{L} \dar{F\boxtimes \phi} \arrow[equal]{u} \\
    \bm^*\gcs{L} \arrow{rr}{\mu}
    && \LxL\\
  \end{tikzcd}
  \]
\end{enumerate}
Morphisms in $\CS(G)$ are defined in the natural way:
\begin{enumerate}
\labitem{(CS.3)}{CS.3} if $\cs{L} = (\gcs{L},\mu,\phi)$ and
  $\cs{L'} = (\gcs{L'},\mu',\phi')$ are character sheaves on $G$ then
  $\Hom_{\CS(G)}(\cs{L},\cs{L'})$ is the set of morphisms $\varphi : \gcs{L} \to \gcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that the following diagrams both commute.
  \[
  \begin{tikzcd}[column sep=40]
  \Frob{G}^* \gcs{L} \rar{\Frob{G}^* \varphi} \arrow[swap]{d}{F} & \Frob{G}^* \gcs{L'} \dar{F'}
  & \arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}} & \arrow[swap]{d}{\mu} m^* \gcs{L} \rar{m^* \varphi} & m^* \gcs{L'} \dar{\mu'} \\
  \gcs{L} \rar{\varphi} & \gcs{L'}
  & {} & \LxL \rar{\tight{1}{\varphi\boxtimes \varphi}{1}} & \gcs{L'} \tight{0}{\boxtimes}{0} \gcs{L'}
  \end{tikzcd}
  \]
\end{enumerate}

%\begin{remark} \label{rem:CS-action}
The rule $(\gcs{L},\mu,\phi) \mapsto (\gcs{L},\phi)$ defines a forgetful functor from character sheaves on $G$ to $\ell$-adic Weil sheaves on $G$ \cite{Deligne:Weil2}*{Def.~1.1.10 (i)}.
As remarked after \cite{Deligne:Weil2}*{Def.~1.1.10}, a $\ell$-adic Weil sheaf $(\gcs{L},\phi)$ on $G$ may be interpreted as a constructible $\ell$-adic sheaf $\gcs{L}$ on $\bG$ together with an action of the Weil group $\Weil{\Fq}$ on $\gcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ on $\bG$. Since we will need this point of view later, we review that interpretation here. Setting 
\[
\phi(\Frob{\Fq}^n) \ceq  \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi)
\]
defines, for every $w\in \Weil{\Fq}$, an isomorphism
\[
\phi(w) : w_G^* \gcs{L}\to \gcs{L},
\]
such that
\[
\phi(uv) = \phi(v) \circ v_G^* \phi(u), \qquad \forall u,v\in \Weil{\Fq}.
\]
Note that $\phi = \phi(\Frob{\Fq})$. 
If we further define $\phi_1(w) \ceq (w_G)_*(\phi(w)^{-1})$ then
\[
\phi_1(w) : (w_G)_* \gcs{L}\to \gcs{L},
\]
is an isomorphism and
\[
\phi_1(uv) = \phi_1(u) \circ (u_G)_* \phi_1(v), \qquad \forall u,v\in \Weil{\Fq},
\]
which means the pair $(\gcs{L},\phi_1)$ is almost an action $\Weil{\Fq}$ on $\gcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ in the sense of \cite{SGA4.5}*{Expos\'e XIII,~1.1}. The pair $(\gcs{L},\phi_1)$ fails to fulfill the requirements of \cite{SGA4.5}*{Expos\'e XIII,~1.1} only because $\Weil{\Fq}$ is profinite. This fact goes to the heart of the notion of a Weil sheaf and is one of the essential aspects of our definition of $\CS(G)$.
%\end{remark}

@@

%\begin{proposition}\label{tensor}
The category $\CS(G)$ of character sheaves on $G$ is a rigid monoidal
category [insert reference] under
$\cs{L} \otimes \cs{L'} \ceq (\gcs{L}\otimes\gcs{L'}, \mu\otimes\mu', \phi\otimes \phi')$
with duals given by applying the functor 
\Clifton{Should this be $\sheafHom(\ - \ ,\EE)$?} $\sheafHom(\ - \ ,(\EE)_{\bG})$ (sheaf hom) to
conditions \ref{CS.0}, \ref{CS.1} and \ref{CS.2}.
%\end{proposition}
(The category of characters sheaves on $G$ is not abelian, so $\CS(G)$
is not a tensor category in the sense of \cite{Deligne:tensorielles}.)
%\begin{proof}\end{proof}
%
This rigid monoidal category structure for $\CS(G)$ puts a group
structure on the set $\CSiso{G}$ of isomorphism classes in
$\CS(G)$; we refer to this as the \cdef{Picard group of character
sheaves} on $G$.\footnote{See {\tt http://ncatlab.org/nlab/show/Picard+group}.}
The main result in this paper is the determination of the Picard group of character sheaves on $G$, when the geometric component group of $G$ is finitely generated.

@@

To define the category of finite character sheaves we must work a bit harder.
Consider the following category $C(G)$:
an object in $C(G)$ is a pair $(f,\psi)$ where: 
$f : H\to G$ an etale cover and a morphism of commutative group $\Fq$-schemes (an isogeny) 
for which the action of $\Gal(\bFq/\Fq)$ on the etale group scheme $\ker f$ is trivial, 
and $\psi : \ker f\to \Aut(V)$ is a representation on a finite-dimensional $\EE$-vector space $V$;
a map $(f,\psi) \to (f',\psi')$ in $C(G)$ is a pair $(\alpha,T)$ 
where $\alpha$ is a morphism of schemes such that $f' = f\circ \alpha$
and $T : V\to V'$ is a linear transformation equivariant 
for the action of $\ker f'$ on $V'$ by $\psi'$ 
and the action of $\ker f'$ on $V$ by $\psi \circ h$.

@@

We say that $(\alpha,T)$ in $C(G)$ is a weak isomorphism if $T$ is an isomorphism of vector spaces. 
Let $C(G)[W^{-1}]$ be the localization of $C(G)$ at weak isomorphisms.
Now, let $C_1(G)$ be the subcategory of $C(G)$ such that $\psi$ is a one-dimensional representation
and let $\CSf(G)$ be the localisation of $C_1(G)$ at weak isomorphisms. 
This is the rigid symmetric monoidal category of \cdef{finite character sheaves} on $G$.

@@

\begin{proof}
 To prove the proposition we exhibit a full and faithful monoidal functor $L_G : \CSf(G) \to \CSb(G)$.
 %
 We will do this by constructing a faithful monoidal functor $L : C_1(G) \to \CSb(G)$ 
 such that $L(\alpha,T)$ is an isomorphism precisely when $(h,T)$ is a weak isomorphism. 
 Factoring along $C_1(G) \to \CSb(G)$ will produce the full and faithful monoidal functor we seek.

 Now we introduce a faithful functor $L : C(G) \to \Loc(G)$ to the category of local systems on $G$.
 For any object $(f,\psi)$ in $C(G)$, let $L(f,\psi)$ be the $\psi$-isotypic component of $f_* V_H$, 
 where $V_H$ refers to $V$ as a constant sheaf on $H$. 
 We may identify the stalk $(f_* V_H)_{\bar g}$ of $f_* V_H$ at a geometric point ${\bar g}$ on $G$ 
 with the $\EE$-vector space $\Hom_\text{set}(\abs{f^{-1}({\bar g})},V)$ 
 and the stalk $L(f,\psi)_{\bar g}$ of $L(f,\psi)$ at a geometric point ${\bar g}$ on $G$ 
 with the $\EE$-vector space consisting of those functions $s: \abs{f^{-1}({\bar s})} \to V$ such that 
 $s(a\cdot {\bar h}) = \psi(a)(s({\bar h}))$ for all $a\in \ker f$ and for all ${\bar h} \in f^{-1}({\bar g})$.
 Since $L(f,\psi)$ is trivialized by $f$, it is a local system on $G$.
 For any map $(\alpha,T)$ in $C(G)$, let $L(\alpha,T) : L(f,\psi) \to L(f',\psi')$ be the homomorphism of local systems defined 
 %at geometric points $L(\alpha,T)_{\bar g} : L(f,\psi)_{\bar g} \to L(f',\psi')_{\bar g}$ 
 by $s \mapsto s' \ceq T\circ s\circ \alpha$. 
 This defines a monoidal functor $L$ from $C(G)$ to $\Loc(G)$. 
 This functor commutes with pull-back along any morphism of group schemes.

 Now consider the localization $C(G)[W^{-1}]$ of $C(G)$ at weak isomorphisms.
 Morphisms in $C(G)[W^{-1}]$ are simple to describe because category $C(G)$ 
 admits a calculus of left fractions\footnote{See {\tt http://ncatlab.org/nlab/show/calculus+of+fractions}.} for weak isomorphisms: 
 every isomorphism in $C(G)$ is a weak isomorphism; 
 push-outs exist in $C(G)$ and the push-out of a weak isomorphism is a weak isomorphism; 
 and weak isomorphisms enjoy the left-cancellation property. 
 It follows that a morphism in $C(G)[W^{-1}]$ is the equivalence class of 
  \[
   \begin{tikzcd}
   (f,\psi) \arrow{r}{(\alpha_1,T_1)} & (f_1,\psi_1) & \arrow[swap]{l}{(\alpha'_1,T'_1)} (f',\psi')
   \end{tikzcd}
  \]
 where $(\alpha',T') : (f,\psi) \to (f',T')$ is any map in $C(G)$ 
 and $(\alpha'_1,T'_1) : (f_1,\psi_1) \to (f',\psi')$ is a weak isomorphism.
% this morphism is commonly denoted by $(\alpha_1,T_1)^{-1}\circ (\alpha'_1,T'_1)$. 
 %The equivalence relation identifies $(\alpha_1,T_1)^{-1}\circ (\alpha'_1,T'_1)$ 
 % with $(\alpha_2,T_2)^{-1}\circ (\alpha'_2,T'_2)$ if there are maps 
 %$(f_1,\psi_1)\to (f_0,\psi_0)$ and $(f_2,\psi_2)\to (f_0,\psi_0)$ making the following diagram commute
 %\[
 %  \begin{tikzcd}
 %  & (f_1,\psi_1) \arrow{d} & \\
 %  (f,\psi) \arrow{ur}{(\alpha_1,T_1)} \arrow{dr}{(\alpha_2,T_2)} & (f_0,\psi_0)  & \arrow[swap]{ld}{(\alpha'_2,T'_2)} \arrow[swap]{lu}{(\alpha'_1,T'_1)} (f',\psi')\\
 %  & (f_2,\psi_2) \arrow{u} & \\
 %  \end{tikzcd}
 %\]
 %where $(f',\psi') \to (f_0,\psi_0)$ is a weak isomorphism.

 Observe that $L$ takes weak isomorphisms in $C(G)$ to isomorphisms in $\Loc(G)$. 
 To see this, let $(h,T) : (f,\psi)\to (f',\psi')$ be a weak isomorphism, 
 where $f:H\to G$ and $f':H'\to G$ as above. 
 Since $T$ is an isomorphism, we may replace $V'$ with $V$ and $T$ with $\id_V$, without loss of generality. 
 Then the equivariance condition on $T$ becomes $\psi' = \psi\circ h$ on the finite etale group scheme $\ker f'$. 
 Since $f' = f \circ \alpha$ and $f$ and $f'$ are etale covers, it follows that $\alpha$ is also an etale cover. 
 It follows that $\alpha : H'\to H$ admits a section $\alpha' : H \to H'$, 
 which provides the inverse to $L(\alpha,T) : L(f,\psi) \to L(f',\psi')$, above. 
 \Clifton{Is it clear?}
 It is also clear that if $L(\alpha,T)$ is an isomorphism of local systems, 
 then $T$ must be an isomorphism of vector spaces, 
 whence $(\alpha,T)$ is a weak isomorphism: $W$ is saturated.

@@

  To produce the isomorphism between $\CSiso{G(\bFq)}$ and
  $\Hh^1(\Weil{\Fq}, G(\bFq)^*)$ we must pick a global section of the line
  bundle $(\gcs{L},\mu)$, for each character sheaf $\cs{L}$ on $G$; 
  such a section exists by Lemma~\ref{lemma:section}.
  We will show that the group isomorphism does not depend on these
  choices.
%
  We use the section $s_\cs{L}$ to define a cocycle $z_\cs{L} \in Z^1(\Weil{\Fq},G(\bFq)^*)$
  as follows. Recall that $\cs{L}$ also determines a family of linear transformations
  $\phi_x : \gcs{L}_{\Frob{\Fq}(x)} \to \gcs{L}_x$ of one-dimensional $\EE$-vector
  spaces, so $\phi_x(s(\Frob{\Fq}(x)))$ is a scalar multiply of $s(x)$, for
  each $x\in G(\bFq)$; let $\tau_\cs{L}(\Frob{\Fq})(x)$ be that scalar, so
  \begin{equation}\label{t}
    \forall x\in G(\bFq),\qquad \phi_{x}(s(\Frob{\Fq}(x)) = \tau_\cs{L}(\Frob{\Fq})(x) s(x).
  \end{equation}
  Conditions \ref{cs.0} and \ref{cs.2}, together with \eqref{section}, guarantee that
  the scalar $z_\cs{L}(\Frob)(x)$ is non-zero, while condition~\ref{cs.1} forces
  $\tau_\cs{L}(\Frob{\Fq})(x+y) = \tau_\cs{L}(\Frob{\Fq})(x) \ \tau_\cs{L}(\Frob{\Fq})(y)$,
  for all $x,y \in G(\bFq)$.
  % \Clifton{The alternative to this approach is to change the
  %   definition of character sheaves to incorporate the notion of
  %   Galois sheaves. That might not be so bad, but would add another
  %   layer of complication, which we wouldn't use profitably, expect
  %   here.}

@@

\begin{lemma}\label{lemma:section}
 If $\cs{L}$ is a character sheaf then there is a non-trivial section of $\gcs{L}$ 
 compatible with $\mu$ in the sense that 
 \begin{equation}\label{section}
  \forall x,y \in G(\bFq), \qquad \mu_{x,y}(s(x+y)) = s(x)\otimes s(y).
 \end{equation}
\end{lemma}

@@

\bib{SGA4.5}{book}{
   author={Deligne, P.},
   title={Cohomologie \'etale},
   series={Lecture Notes in Mathematics, Vol. 569},
   note={S\'eminaire de G\'eom\'etrie Alg\'ebrique du Bois-Marie 1963-64 SGA
   4$\frac{1}{2}$; 
   Avec la collaboration de J. F. Boutot, A. Grothendieck, L. Illusie et J.
   L. Verdier},
   publisher={Springer-Verlag},
   place={Berlin},
%date={4h},
   date={1977},
%   pages={iv+312pp},
%   review={\MR{0463174 (57 \#3132)}},
}

\bib{SGA7.2}{book}{
   title={Groupes de monodromie en g\'eom\'etrie alg\'ebrique. II},
%   language={French},
 author={Deligne, P.},
 author={Katz, N.},
   series={Lecture Notes in Mathematics, Vol. 340},
   note={S\'eminaire de G\'eom\'etrie Alg\'ebrique du Bois-Marie 1967--1969
   (SGA 7 II);
   Dirig\'e par P. Deligne et N. Katz},
   publisher={Springer-Verlag},
   place={Berlin},
   date={1973},
%   pages={x+438},
%   review={\MR{0354657 (50 \#7135)}},
}
